---
title: "MFtools README"
author: "Dan Puddephatt"
date: "`r Sys.Date()`"
output:
    github_document
---
The MFtools library is used for investigating the input and output files from MODFLOW simulations. 
The library is based on fixed-width file formats as are generated when using Groundwater Vistas. 
This library has been tested using Groundwater Vistas files. The library includes functions for the following
tasks:

* Reading the discritization file (readdis(rootname))
* Reading the layer property flow file (readlpf(rootname))
* Reading the basic transport package file (readbtn(rootname))
* Reading the headsave file (readhds(rootname, NLAY, NTTS))
* Reading the unformatted concentration file (readucn(ucrootname, NLAY, NTTS))
* Reading no-flow array file (readnoflow(FILE, rootname))
* Writing the unformatted concentration file based on a concentration dataframe (writeucn(ucn, ofl))

The library also includes a utility function for calculating contaminant mass based on information from the
headsave file (i.e., saturated thickness), porosity read from the basic transport package,
 and concentrations from the unformatted concentration file (masscalc(BTN, HDS, UCN))

The functions from the MFtools library use rootnames to read files. Rootnames
are based on the MODFLOW rootname convention such that files consist of
a rootname and an extension that describes the MODFLOW file.

(NOTE that by my personal convention my script headings include
the libraries that I will be using):

```{r echo = FALSE}
ROOT_NM <- system.file("extdata", "Ogll_sv2.01_ss.dis", package = "MFtools") %>% 
tools::file_path_sans_ext()
```

```{r LOAD, message = FALSE}
library(MFtools)
library(tidyverse)
library(magrittr)
library(scales)

rnm <- ROOT_NM
```

With the rootname we can now read in the discretization file

```{r DIS}
d <- readdis(rnm)
d
```

The structure of the variable, d, is a list consisting of the information
from the MODFLOW discretization file. For indexing purposes, some of the 
information is arranged in a dataframe, such as the TOP and BOT variables that
describe the top and bottom elevations of each model layer.

MFtools includes a function to read the layer property flow file (.lpf).

```{r LPF}
p <- readlpf(rnm)
p
```

The .lpf file has lots of really useful information. We can isolate this information using the
"$" indexing operation.

```{r PROPS}
p$PROPS

p$PROPS %>% group_by(LAY) %>% summarise(MIN = min(HK), MEDIAN = median(HK), MAX = max(HK))
```

The model is a four layer model. Summarising does not give us a lot of
information about the distribution of hydraulic conductivity values.

We can, instead, use ggplot to plot a histogram of hydraulic conductivity values
by layer

```{r HISTO_K}
p$PROPS %>% ggplot(aes(x = HK)) +
            geom_histogram(aes(y = (..count..)/tapply(..count.., ..PANEL.., sum)[..PANEL..]),
                colour = "black", 
                fill = "yellow", 
                alpha = 0.6) +
            scale_y_continuous(limits = c(0, 0.2), 
                breaks = seq(0, 0.2, 0.05), 
                labels = percent, 
                name = "FREQUENCY COUNT (PERCENT)") +   
            scale_x_log10(breaks = 10^(-10:10), 
                 minor_breaks = c(1:9 %o% 10^(-10:10)), 
                 labels = scales::trans_format("log10", math_format(10^.x))) +
            facet_wrap(~LAY, ncol = 1) +
            labs(x = "HYDRAULIC CONDUCTIVITY (FEET / DAY)") +
            theme_bw()
```

We can also investigate groundwater elevations by model layer

```{r GWE}
h <- readhds(rnm, NLAY = d$NLAY, 1)
h %>% filter(GWE > 2000) %>% filter(GWE < 10000) %>%
            ggplot(aes(x = GWE)) +
            geom_histogram(aes(y = (..count..)/tapply(..count.., ..PANEL.., sum)[..PANEL..]),
                colour = "black", 
                fill = "blue", 
                alpha = 0.6) +
            scale_y_continuous(limits = c(0, 0.2), 
                breaks = seq(0, 0.2, 0.05), 
                labels = percent, 
                name = "FREQUENCY COUNT (PERCENT)") +   
            facet_wrap(~LAY, ncol = 1) +
            labs(x = "GROUNDWATER ELEVATION (FEET AMSL)") +
            theme_bw()
```
            